---
title: "Assignment-2:- The Transportation Model"
author: "ANURODH-SINGH"
date: "2023-10-02"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


**Summary**

***

1.The Northern Airplane Company used linear programming (LP) to optimize jet airplane          production costs, resulting in a minimum cost of $77.3 million. 
2.The LP model included 20 decision variables and 9 constraints, but violated the fundamental   assumption of the transportation problem that supply must equal demand. 
3.To address this, a dummy demand for 30 jet airplanes was introduced.
4.The LP model was solved using the "read.lp," "solve," and "get.objective" functions. 
5.The decision variables were represented as "X(ij)," with the condition "i<=j" applied to     ensure that 'i' is always less than or equal to 'j.'
6.Ultimately, the schedule developed for the production of engines in each of the four months   can be utilized by the production manager.
7.The notations (Xij) signify the number of planes produced in month 'i' for month 'j,' with   specific values assigned as follows: X11=10, X12=15, X23=5, X33=20, X34=10, and X44=10.
8.This production schedule can be used by the production manager to optimize jet airplane      production.



```{r}
library(lpSolveAPI)
```

**Making an Lp file and calling to read it:-**
```{r}
x <- read.lp("C:/Users/ASUS/Desktop/QMM-ASSIGNMENT/classwork/12-oct-QMM.lp")
write.lp(x,"test.out")
x
```

**solved the lp model**
```{r}
solve(x)
get.objective(x)
get.variables(x)
get.constraints(x)
get.sensitivity.objex(x)
get.sensitivity.rhs(x)
```

